WORCS: A Workflow for Open Reproducible Code in Science

### 2020-10-31 paper/Manuscript_DS.Rmd
RStudio public benefit, encourage porting to other languages

The `R` package `worcs` automates much of the workflow
for `R` users.
Several arguments support the choice to implement this workflow in `R`.
First, `R` and all of it extensions are free open source software, which
make it a tool of choice for open science.
Second, all tools required for an open science workflow are implemented in `R`,
most of these tools are directly accessible through the RStudio user interface,
and some of them are actively developed by the team behind RStudio.
Third, `R` is the second-most cited statistical software package [@muenchenPopularityDataScience2012], following SPSS, which has no support for any of the open science tools discussed in this paper.
Fourth, `R` is well-supported by
a vibrant and inclusive online community,
which develops new methods and packages,
and provides support and tutorials for existing ones.
Finally, `R` is highly interoperable:
Packages are available to read and write nearly every filetype,
including DOCX, PDF (in APA style), and HTML. 
Moreover, wrappers are available for many tools developed in other programming languages, and code written in other programming languages can be evaluated from `R`, including `C++`, 
`Fortran`, and `Python` [@allaireMarkdownPythonEngine2020]. There are excellent free
resources for learning `R` [e.g., @grolemundDataScience2017].

>>>>

WORCS is, first and foremost, a conceptual workflow that could be implemented in any software environment.
As of this writing, the workflow has been implemented for `R` users in the package `worcs` [@worcspackage2020].
Several arguments support the choice to implement this workflow in `R` first, although we encourage developers to port the software to other languages.
First, `R` and all of it extensions are free open source software, which
make it a tool of choice for open science.
Second, all tools required for an open science workflow are implemented in `R`,
most of these tools are directly accessible through the RStudio user interface,
and some of them are actively developed by the team behind RStudio.
Third, unlike any competitor, RStudio is a Public Benefit Corporation, whose primary purpose is explicitly aligned with open science principles:

> RStudio’s primary purpose is to create free and open-source software for data science, scientific research, and technical communication. This allows anyone with access to a computer to participate freely in a global economy that rewards data literacy; enhances the production and consumption of knowledge; and facilitates collaboration and reproducible research in science, education and industry.

Fourth, `R` is the second-most cited statistical software package [@muenchenPopularityDataScience2012], following SPSS, which has no support for any of the open science tools discussed in this paper.
Fifth, `R` is well-supported by
a vibrant and inclusive online community,
which develops new methods and packages,
and provides support and tutorials for existing ones.
Finally, `R` is highly interoperable:
Packages are available to read and write nearly every filetype,
including DOCX, PDF (in APA style), and HTML. 
Moreover, wrappers are available for many tools developed in other programming languages, and code written in other programming languages can be evaluated from `R`, including `C++`, 
`Fortran`, and `Python` [@allaireMarkdownPythonEngine2020]. There are excellent free
resources for learning `R` [e.g., @grolemundDataScience2017].
